---
title: "Missing Data"
output: word_document
date: "2024-09-22"
---
Libraries
```{r}
library(tidyverse)
library(VIM)
library(mice)

```
read in dataset
```{r}
Grades = read.csv("class-grades.csv")
```

**Question 1** 
For which variable is there missing data? 
```{r}
str(Grades)
summary(Grades)
skim(Grades)
```
***ANSWER*** 
Tutorial, Midterm, TakeHome, Final

**Question 2**
Use the VIM package to visualize missingness. Are there any students that are missing multiple piecies of data? 
```{r}
vim_plot = aggr(Grades, numbers = TRUE, prop = c(TRUE, FALSE),cex.axis=.7)
#the cex.axis reduces size of text on x-axis so labels fit better
```
**Answer** 
Yes, 1 student

**Question 3**
Use row-wise deletion of missing values to create a new data frame. How many rows remain in this data frame? 
```{r}
Grades_rowdel = Grades %>% drop_na() 
#alternatively can specify which variable(s) on which to do the drop_na
#drop_na(Age)
skim(Grades_rowdel)
```
**Answer**
89 Rows

**Question 4**
Use column-wise deletion of missing values to create a new data frame (from the origninal data frame not from the data frame create in Task 3) How many columns remain in this data frame? 
```{r}
Grades_coldel = Grades %>% select(-Tutorial, -Midterm, -TakeHome, -Final) 
skim(Grades_coldel)
```
**Answer**
0, None

**Question 5**
Which approach (Taks 3 or Task 4) seems preferable for this dataset? Briefly discuss your answer.
# row-wise deletion would be the best method because if you use column-wise deletion you will be left with no data to use due to all of the columns containing missing data. 

**Task 6**
Use the "mice" package to impute the missing data in this dataset. Set "m=5" and "method =" pmm" in imputation. You MUST use a random number see of 123 for this imputation. What is the mean value for the "Final" variable after imputation? 
```{r}
set.seed(123) #sets seed for random number generator
imp_missing = mice(Grades, m=5, method='pmm', printFlag=FALSE)
#m is the number of imputations, 5 is a reasonable value as a default
#pmm is "predictive mean matching" = imputation method for numeric data
#printFlag reduces amount of output
summary(imp_missing)
```
#Merge the imputed values into our Grades data frame  
```{r}
Grades_complete = complete(imp_missing) 
summary(Grades_complete)
```
***Answer***
68.33
